---
title: "Empirical and Simulated Anolis Ensembles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Empirical and Simulated Anolis Ensembles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(TrajectoryTrees)
```

The source package includes the release Anolis inputs: 500 empirical stochastic
maps and 1,000 simulated maps. `load_anolis_trees()` gives these collections
stable names and supports indexed subsets.

```{r load}
empirical_maps <- load_anolis_trees("empirical", limit = 2)
simulated_maps <- load_anolis_trees("simulated", limit = 2)

length(empirical_maps)
length(simulated_maps)
```

## Constructing both groups

Nested input produces a nested result with the same names.

```{r construct}
trajectories <- make_trajectory_objects(
  list(
    empirical = empirical_maps,
    simulated = simulated_maps
  )
)

names(trajectories)
```

For the complete 1,500-map analysis, use batches:

```{r full-batches, eval=FALSE}
empirical_manifest <- make_trajectory_batches(
  load_anolis_trees("empirical"),
  output_dir = "anolis_batches/empirical"
)

simulated_manifest <- make_trajectory_batches(
  load_anolis_trees("simulated"),
  output_dir = "anolis_batches/simulated"
)
```

## Posterior distributions

Each trajectory list becomes one posterior distribution. The C++ reducer is
exact and uses bounded working blocks.

```{r distributions}
distributions <- make_trajectory_distributions(
  trajectories,
  include = c("P", "T", "SS", "TT"),
  backend = "cpp"
)

distributions$empirical
distributions$simulated
```

The distribution object keeps posterior means, medians, lower and upper
quantiles, and the number of observed values for each numeric cell.

```{r plot-distribution, fig.width=7, fig.height=4}
plot_trajectory_distribution(
  distributions$empirical,
  surface = "SS",
  title = "Empirical Anolis posterior summary"
)
```

## Comparing empirical and simulated summaries

`compare_trajectory_distributions()` is an exact path-by-path diagnostic when
the two distributions have compatible numeric surfaces.

```{r compare}
comparison <- compare_trajectory_distributions(
  distributions$empirical,
  distributions$simulated,
  surfaces = c("SS", "TT")
)

head(comparison)
```

The comparison reports structure mismatches as well as maximum absolute
numeric differences, so results can be filtered to the biological surfaces of
interest without relying on implementation-specific object names.
