---
title: "Constructing and Summarizing Trajectory Trees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Constructing and Summarizing Trajectory Trees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(TrajectoryTrees)
```

TrajectoryTrees converts a stochastic character map into a stable
`trajectory_tree`. The package keeps the implementation private: user code uses
the same names even when the internal engine changes.

## One stochastic map

The package includes a small deterministic map for examples and tests.

```{r construct}
simmap <- readRDS(system.file(
  "extdata", "generated_000001.rds",
  package = "TrajectoryTrees"
))

trajectory <- make_trajectory_objects(simmap)
trajectory
```

The returned object contains the input mapped phylogeny, scenario and transition
trees, summary statistics, through-time matrices, scenario coordinates, and
the resolved root policy.

```{r inspect}
names(trajectory)
names(trajectory_summary(trajectory))
names(trajectory_through_time(trajectory))
```

## Lists and nested lists

Names and input nesting are retained. This makes posterior samples and
empirical/simulation groupings use the same constructor.

```{r lists}
posterior <- make_trajectory_objects(
  list(first = simmap, second = simmap),
  include_tt = FALSE
)

groups <- make_trajectory_objects(
  list(
    empirical = list(first = simmap),
    simulated = list(first = simmap)
  ),
  include_tt = FALSE
)

names(posterior)
names(groups)
```

## Keeping only needed fields

Large posterior jobs often need fewer surfaces than an interactive inspection.

```{r select}
compact <- select_trajectory_fields(
  trajectory,
  include = c("phylo", "scenario", "trans", "ss"),
  ss_include = c("tree_wide", "by_state", "by_path")
)

names(compact)
names(compact$ss)
```

## Through-time plots

```{r plot-tt, fig.width=7, fig.height=4}
plot_trajectory_through_time(
  trajectory,
  metric = "ltt",
  component = "state",
  title = "Lineages through time by state"
)
```

Tree plots use `phytools::plotSimmap()` when phytools is installed and otherwise
fall back to `ape::plot.phylo()`.

```{r plot-tree, fig.width=7, fig.height=5}
plot_trajectory_tree(trajectory, tree = "phylo", fsize = 0.7)
```

## Resumable batches

For large ensembles, construct directly into batch files:

```{r batches, eval=FALSE}
manifest <- make_trajectory_batches(
  simmaps,
  output_dir = "trajectory_batches",
  batch_size = "auto"
)

trajectories <- load_trajectory_batches(manifest)
```

The manifest records input identity, shape, names, construction controls, and
completed files. `resume = TRUE` continues only a compatible incomplete run.
